\vspace{1.0cm} \Large \bf Integrating influenza antigenic dynamics with molecular evolution \\ or \\ Cartographic characterization of the antigenic, evolutionary and competitive dynamics of seasonal influenza viruses

### 2013-02-04 flux.tex
Implement wording fix suggested by @evogytis.

Interestingly, although antigenic clusters seem to emerge in a punctuated fashion in influenza A/H3N2, year-to-year antigenic drift in the mean antigenic phenotype of the virus population is relatively constant through time with most years exhibiting an intermediate level of drift (Figure~\ref{jumps}).
We calculate year-to-year antigenic drift for years 1992 to 2011 by calculating the average location along dimension 1 of phylogenetic lineages present in the tree at year $i$ and comparing this location to the average location of phylogenetic lineages present in the tree at year $i-1$.
For A/H3N2, the largest cluster transition we observe is the emergence of the Sydney/97 cluster, evident as the appearance in 1997 of a large gap in virus locations along antigenic dimension 1 (Figure~\ref{drift}).
However, in 1997 only three of the nine viruses sampled show the Sydney/97 phenotype, while the remaining six viruses exhibit the old Wuhan/95 phenotype.
In 1998, three of the four viruses sampled show the Sydney/97 phenotype, with the remaining virus still showing the Wuhan/95 phenotype.
In 1999, all sampled viruses exhibit the Sydney/97 phenotype.
Thus, we observe fairly equal rates of year-to-year drift in mean antigenic phenotype between 1996 and 1999, rather than all drift in mean antigenic phenotype coinciding with the appearence of a new antigenic variant in 1997 (Figure~\ref{jumps}).
It is interesting that this phenomenon occurs despite the fact that antigenic drift is accomplished, not by steady movement of the entire virus population, but by the successive emergence of drifted variants that each come to take over the virus population.
Because drifted variants require, on average, 2--3 years to take over the virus population, rates of year-to-year drift do not always spike heavily at the appearance of a new variant.

>>>>

Interestingly, although antigenic clusters seem to emerge in a punctuated fashion in influenza A/H3N2, year-to-year antigenic drift in the mean antigenic phenotype of the virus population is relatively constant through time with most years exhibiting an intermediate level of drift (Figure~\ref{jumps}).
We calculate year-to-year antigenic drift for years 1992 to 2011 by calculating the average location along dimension 1 of phylogenetic lineages present in the tree at year $i$ and comparing this location to the average location of phylogenetic lineages present in the tree at year $i-1$.
For A/H3N2, the largest cluster transition we observe is the emergence of the Sydney/97 cluster, evident as the appearance in 1997 of a large gap in virus locations along antigenic dimension 1 (Figure~\ref{drift}).
However, in 1997 only three of the nine viruses sampled show the Sydney/97 phenotype, while the remaining six viruses exhibit the old Wuhan/95 phenotype.
In 1998, three of the four viruses sampled show the Sydney/97 phenotype, with the remaining virus still showing the Wuhan/95 phenotype.
In 1999, all sampled viruses exhibit the Sydney/97 phenotype.
Thus, we observe fairly equal rates of year-to-year drift in mean antigenic phenotype between 1996 and 1999, rather than all drift in mean antigenic phenotype coinciding with the appearence of a new antigenic variant in 1997 (Figure~\ref{jumps}).
It is interesting that this phenomenon occurs despite the fact that antigenic drift is accomplished, not by steady movement of the entire virus population, but by the successive emergence of drifted variants that sweep through the virus population.
Because drifted variants require, on average, 2--3 years to take over the virus population, rates of year-to-year drift do not always spike heavily at the appearance of a new variant.
